---
title: "data_prep"
output: html_document
---

# Set up
Install Packages
```{r}
list.of.packages <- c("data.table", "dplyr")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

Load packages
```{r warning=FALSE}
library("data.table")
library("dplyr")
```

Load data
```{r}
data <- fread('../../data/parsed_playlists.csv', sep='\t', quote = '')
nrow1 <- nrow(data)
nrow1
```

# 4. Data
## 4.1 Sample collection

Remove duplicate rows
```{r}
data <- unique(data)
nrow2 <- nrow(data)
nrow2

#Number of rows removed because they were duplicates:
n_duplicates_removed <- nrow1-nrow2
print(n_duplicates_removed)
```

Change values to dates/numeric
```{r warning=FALSE}
data[, added_at := as.Date(added_at)]
data[removed_at == "None", removed_at := NA]
data[, removed_at := as.Date(removed_at)]
data[, release_date_first := as.Date(release_date_first)]
data[, release_date_all := as.Date(release_date_all)]

data[, song_duration_ms := as.numeric(song_duration_ms)]
data[, track_spotify_popularity := as.numeric(track_spotify_popularity)]
data[, pl_followers := as.numeric(pl_followers)]
data[, num_track_in_pl := as.numeric(num_track_in_pl)]
data[, pl_fdiff_month := as.numeric(pl_fdiff_month)]
data[, pl_fdiff_week := as.numeric(pl_fdiff_week)]
data[, pl_monthly_listeners := as.numeric(pl_monthly_listeners)]
data[, pl_active_ratio := as.numeric(pl_active_ratio)]

```

Empty, "None", "," cells to NA
```{r}
data$pl_owner <- ifelse(data$pl_owner=="", NA, data$pl_owner)
data$album_id[data$album_id=="None"] <- NA
data$album_name[data$album_name==""] <- NA
data$album_name[data$album_name==","] <- NA
data$first_artistname[data$first_artistname==""] <- NA
data$pl_tags_genre[data$pl_tags_genre==""] <- NA
data$pl_name[data$pl_name==""] <- NA
data$pl_name[data$pl_name==","] <- NA

#Change missing album ids to album names
data$album_id <- ifelse(is.na(data$album_id), data$album_name, data$album_id)
```

## 4.2 Data preparation and variable operationalization
### 4.2.1 Tracks

Track name "Home":
```{r}
data %>% filter(track_name=="home") %>% distinct(track_id, .keep_all=T) %>% 
  distinct(artist_id) %>% nrow()
```


Only keep tracks with release dates in 2017
(we collected data for only these tracks)
```{r}
tracks_before <- data %>% distinct(track_id) %>% nrow()
data <- data %>% filter(release_date_first>"2016-12-31" & release_date_first<"2018-01-01")

# N tracks removed that were not released in 2017 or
# for which release date was unknown
tracks_after <- data %>% distinct(track_id) %>% nrow()
n_removed <- tracks_before-tracks_after
```

Indicate whether a track was released by a major label or an indie label
```{r}

labels_warner = c('Warner Music' = 'warner[ ]music|warner[ ]home|warner[ ]special|warner[ ]strategic|warner[.]esp',
			       'Asylum Records' = 'asylum[ ]records|([|]|^)atlantic[ ]records|atlantic[ ]|([|]|^)atlantic([|]|$)|elektra[ ]records|([|]|^)elektra([|]|$)|warner[ ]music[ ]nashville|warner[ ]bros|elektra[ ]nashville',
			       'Big Beats Records' = 'big[ ]beat[ ]|([|]|^)big beat([|]|$)',
             'Canvasback Music' = 'canvasback',
             'Parlophone Label Group' = 'parlophone|FFR[ ]records|([|]|^)FFRR([|]|$)|virgin[ ]classics|emi[ ]classics|[ ]erato[ ]|([|]|^)erato|warner[ ]classics',
             'Reprise Records' = '([|]|^)reprise[ ]',
             'Fueled By Ramen' = 'Fueled[ ]by[ ]ramen',
             'Nonesuch Records' = 'nonesuch[ ]records|([|]|^)nonesuch([|]|$)',
             'Rhino Entertainment' = 'rhino[ ]entertainment|([|]|^)rhino',
             'Roadrunner Records' = '([|]|^)roadrunner',
             'Sire Records' = '([|]|^)sire[ ]records|([|]|^)sire[ ]|([|]|^)sire[ ]([|]|$)',
             'East West' = 'east[ ]west|eastwest',
             'Warner (all combined)' = '([|]|^)warner|[(]warner[)]|asylum[ ]records|big[ ]beat[ ]records|canvasback[ ]music|parlophone[ ]label[ ]group|reprise[ ]records|fueled[ ]by[ ]ramen|nonesuch[ ]records|rhino[ ]entertainment|roadrunner[ ]records|sire[ ]records|east[ ]west')

labels_universal = c('Universal Music Group' = '([|]|^)universal|([|]|^)universal[ ]music[ ]japan|([|]|^)universal[ ]sigma|([|]|^)universal[ ]international|([|]|^)geneon[ ]universal|nbcuniversal|universal[ ]licensing[ ]music|([|]|^)universal[ ]music[ ]|universal[ ]music[ ]spain|universal[ ]m..z.k|([|]|^)universal records|([|]|^)universal[ ]records[ ]|([|]|^)universal[ ]republic[ ]records',
					 'Capital Music Group' = 'capitol|astralwerks|blue[ ]{0,2}note|([|]|^)caroline[ ]|deep[ ]{0,2}well|([|]|^)harvest|([|]|^)metamorphosis|motown|quality[ ]{0,2}control|([|]|^)virgin[ ]|([|]|^)virgin([|]|$)',
                     'Decca Classics' = 'decca|ecm([ ]|$|[|])|([|]|^)mercury|([|]|^)mercury[ ]classics|([|]|^)mercury[ ]records',
                     'Def Jam Recordings' = 'def[ ]{0,2}jam|artium|g.o.o.d|([|]|^)good([|]|$)|good[ ]records',
                     'Deutsche Grammophon' = 'deutsche[ ]grammophon|grammophon',
                     'Eagle Rock Entertainment' = 'eagle[ ]rock|eagle[ ]records', 
                     'EMI' = 'emi[ ]|([|]|^)emi|([|]|^)emi[ ]music|emi[-]',
                     'Interscope' = 'interscope|geffen|A[&]M|([|]|^)222([|]|$)|aftermath|dreamville|insomniac[ ]|kidinakorner|shady[ ]{0,2}records|([|]|^)shady',
                     'Island Records' = '([|]|^)island[ ]records|4th & Broadway|universal[ ]island|([|]|^)island[ ]|([|]|^)island([|]|$)',
                     'Polydor Records' = '([|]|^)polydor[ ]|([|]|^)polydor[ ]([|]|$)|([|]|^)fiction[ ]records|([|]|^)fiction([|]|$)|polydor', 
                     'Republic Records' = '([|]|^)republic[ ]records|universal[ ]republic|([|]|^)american[ ]recordings|([|]|^)Brushfire[ ]records|([|]|^)casablanca[ ]records|([|]|^)cash[ ]money[ ]records|john[ ]varvatos|([|]|^)lava[ ]records|lightworkers|([|]|^)the[ ]voice', 
					           'Republic Records 2' = '([|]|^)republic[ ]records|([|]|^)american[ ]recordings|([|]|^)Brushfire[ ]records|([|]|^)casablanca[ ]records|([|]|^)cash[ ]money[ ]records|john[ ]varvatos|([|]|^)lava[ ]records|lightworkers|([|]|^)the[ ]voice', 
					           'Universal Music Enterprises' = '([|]|^)universal[ ]|([|]|^)universal([|]|$)|T[-]{0,1}boy',
                     'Universal Music Group Nashville' = 'capitol[ ]{0,1}records[ ]{0,1}nashville|emi[ ]{0,1}records[ ]{0,1}nashville|mca[ ]{0,1}nashville|mercury[ ]{0,1}nashville|show[-]{0,1}dog',
                     'Universal Music Latin Entertainment' = 'capitol[ ]{0,1}latin|disa[ ]{0,1}records|fonovisa|machete[ ]{0,1}music|universal[ ]{0,1}music[ ]{0,1}latino',
                     'Verve Label Group' = 'verve[ ]label[ ]group|verve[ ]{0,1}records|decca[ ]{0,1}gold|universal[ ]{0,1}music[ ]{0,1}classics|decca[ ]{0,1}broadway|verve[ ]{0,1}group|([|]|&)verve([|]|$)|verve[ ]{0,1}music[ ]{0,1}group', 
                     'PM:AM Recordings' = 'PM[:]AM|pm[ ]{0,1}am',
                     'Spinefarm Records' = 'spinefarm',
                     'SpinnUp' = 'SpinnUp',
                     'Disques Barclay' = 'disques[ ]{0,1}barclay|([|]|^)barclay',
                     'Varese Sarabande' = 'var.se',
					 'Universal Music (combined)' = 'varese[ ]sarabande|disques[ ]barclay|spinnup|spinefarm[ ]records|pm:am[ ]recordings|verve[ ]label[ ]group|universal[ ]music[ ]latin[ ]entertainment|universal[ ]music[ ]group[ ]nashville|universal[ ]music[ ]enterprises|republic[ ]records|polydor[ ]records|island[ ]records|interscope|emi|eagle[ ]rock[ ]entertainment|deutsche[ ]grammophon|def[ ]jam[ ]recordings|decca[ ]classics|capitol[ ]music[ ]group|universal[ ]music[ ]group')

labels_sony = c('Columbia Records'='CBS[ ]columbia|([|]|^)columbia|hypnotize[ ]minds',
                'Columbia Records 2' = 'dreamville[ ]entertainment|small.*giant|startime[ ]international|blue[ ]propaganda', 
                'RCA Records' = '([|]|^)rca|([|]|^)bystorm.*entertainment|([|]|^)nappy[ ]boy|([|]|^)j[ ]records', 
                'Epic Records' = '([|]|^)epic|([|]|^)battery|([|]|^)freebandz|([|]|^)bad[ ]{0,1}boy[ ]records|([|]|^)volcano|vested[ ]in[ ]culture', 
                'Sony Music Nashville' = 'sony[ ]music|([|]|^)arista|([|]|^)columbia[ ]nashville|rca[ ]records[ ]nashville', 
                'Zomba Music Group' = '([|]|^)zomba|([|]|^)jive[ ]records|([|]|^)verity|([|]|^)silvertone',
                'RED Music Distribution' = '([|]|^)red[ ]music[ ]|odd[ ]future|([|]|^)red[ ]ink|cinematic[ ]music|([|]|^)reach[ ]records', 
                'Legacy Recordings' = '([|]|^)legacy[ ]recordings|([|]|^)laface', 
                'Sony Music Latin' = 'sony.*latin',
                'Ariola Records' = 'ariola',
                'Sony Masterworks' = 'sony[ ]masterworks|([|]|^)bluebird|([|]|^)okeh|portrait[ ]records|([|]|^)portrait|([|]|^)arte[ ]nova|sony[ ]classical|flying[ ]buddha|([|]|^)masterworks', 
                'Provident Label Group' = '([|]|^)provident|essential[ ]records|flicker[ ]records|beach[ ]street|reunion[ ]records|essential[ ]worship',
                'Century Media Records' = 'century[ ]media|([|]|^)century record|people[ ]like[ ]you|insideout[ ]music|superball[ ]music',
				'Sony Music Entertainment' = 'Sony[ ]BMG|([|]|^)BMG|columbia[ ]music|sony[ ]music|Columbia[ ]records|RCA[ ]Records|Epic[ ]Records|Sony[ ]Music[ ]Nashville|Zomba[ ]Music[ ]Group|RED[ ]Music[ ]Distribution[ ]|Legacy[ ]Recordings|Sony[ ]Music[ ]Latin|Ariola[ ]Records|Sony[ ]Masterworks|Provident[ ]Label[ ]Group|Century[ ]Media[ ]Records')

label_iter=c(labels_warner, labels_universal, labels_sony)

data <- data.table(data)

data[, major_label:=0]

for (lbl in label_iter) {
  data[grepl(lbl, album_label, ignore.case=TRUE), major_label:=1]
}

data$album_label[data$album_label=="" | data$album_label==","] <- NA
data$major_label[is.na(data$album_label)] <- NA

data$major_label[data$major_label==1] <- "Major label"
data$major_label[data$major_label==0] <- "Indie label"

```

### 4.2.2 Playlists

Categorize playlist owners
```{r}
data <- data.table(data)
official_pl <- filter(data, pl_official==TRUE)
data[, owner:=0]
data$owner[data$pl_official==TRUE] <- "Spotify"
data[grepl("filtr", pl_owner, ignore.case=TRUE), owner:="Major label"]
data[grepl("topsify", pl_owner, ignore.case=TRUE), owner:="Major label"]
data[grepl("digster", pl_owner, ignore.case=TRUE), owner:="Major label"]
data$owner[data$owner==0] <- "Other"

data[, cur_spotify:=0]
data[, cur_filtr:=0]
data[, cur_topsify:=0]
data[, cur_digster:=0]
data$cur_spotify[data$pl_official==TRUE] <- 1
data[grepl("filtr", pl_owner, ignore.case=TRUE), cur_filtr:=1]
data[grepl("topsify", pl_owner, ignore.case=TRUE), cur_topsify:=1]
data[grepl("digster", pl_owner, ignore.case=TRUE), cur_digster:=1]

```

(4.2.1:) Change all album labels that appear in major label playlists to major label
```{r}
# N albums now marked as indie label added to major label pls
data %>% filter(owner=="Major label" & major_label=="Indie label") %>% 
  distinct(track_id) %>% nrow()

labels_in_major_pl <- data %>% filter(owner=="Major label") %>% distinct(album_label) %>% 
  pull(album_label)

data$major_label[data$album_label %in% labels_in_major_pl] <- "Major label"
```

Playlist genre dummy variables
```{R genres}
no_distinct_playlists <- distinct(data, pl_id, .keep_all=TRUE)

all_tags = strsplit(no_distinct_playlists$pl_tags_genre, ',')
# check out the data
all_tags[3] # or 2, or 3...

freq_table = unlist(all_tags)
freq_table = table(freq_table)

freq_table = freq_table[order(freq_table, decreasing=T)]

top20 = names(freq_table[1:20])

data <- data.table(data)

for (genre in top20) {
  newvariable = gsub('[ ]|[&]|[-]','',genre)
  data[, paste(newvariable):=0]
  data[grepl(genre, pl_tags_genre), paste(newvariable):=1]
  
}
```

Playlists with unknown followers
```{r}
# N playlists for which n followers unknown 
data %>% distinct(pl_id, .keep_all=TRUE) %>% filter(is.na(pl_followers)) %>% nrow()

data <- data %>% filter(!is.na(pl_followers))
```

Remove tracks with playlist addition dates before their release date
```{r}
remove_tracks <- data %>% filter(added_at<release_date_first) %>% 
  distinct(track_id) %>% pull(track_id)
length(remove_tracks)
data <- data %>% filter(!track_id %in% remove_tracks)
```

If a track is added to the same pl multiple times, keep only first addition
```{r}
data <- data %>% arrange(track_id, added_at)
data <- data %>% group_by(track_id) %>% distinct(pl_id, .keep_all=TRUE)
data <- data %>% ungroup()

nrow(data)
```

Playlist popularity above median = 1, below or at median = 0
```{r}
data$pop_above_median <- "Popular"
median_followers <- data %>% distinct(pl_id, .keep_all=T) %>% 
  filter(!is.na(pl_followers)) %>% group_by(owner) %>% 
  summarise(median=median(pl_followers))
spotify_median <- median_followers %>% filter(owner=="Spotify") %>% pull(median)
major_median <- median_followers %>% filter(owner=="Major label") %>% pull(median)
other_median <- median_followers %>% filter(owner=="Other") %>% pull(median)

data$pop_above_median[data$owner=="Spotify" & data$pl_followers<spotify_median] <- "Unpopular"
data$pop_above_median[data$owner=="Major label" & data$pl_followers<major_median] <- "Unpopular"
data$pop_above_median[data$owner=="Other" & data$pl_followers<other_median] <- "Unpopular"
```

t_add
```{r}
first_adds <- data %>% group_by(track_id) %>% summarise(track_first_add = min(added_at))

data <- data %>% left_join(first_adds, by=c("track_id"="track_id"))

data <- data %>% mutate(t_add=added_at-track_first_add)
```

Save as csv
```{r}
write.csv(data, "../../gen/data-preparation/input/data.csv", row.names=FALSE)
```

