---
title: "descriptives"
output: html_document
---

Install Packages
```{r}
list.of.packages <- c("data.table", "dplyr", "tidyr", "ggplot2", "sqldf")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
```

Load packages
```{r}
library("data.table")
library("ggplot2")
library("dplyr")
library("sqldf")
library("tidyr")
```

Load data
```{r}
data <- read.csv("../../gen/data-preparation/input/data.csv", header = TRUE)
```

## 4.3 Descriptive statistics

### 4.3.1 Tracks

Total number of playlist additions
```{r}
# Total number of playlist additions in data
nrow(data)

# Number of tracks in data
distinct_tracks <- data %>% distinct(track_id, .keep_all=TRUE)
nrow(distinct_tracks)
```

Number of playlist additions per track
```{r}
data %>% count(track_id) %>% summarise(min(n), max(n), mean(n), sd(n))

# Plot distribution of number of playlist additions per track.
data %>% count(track_id) %>% filter(n<101) %>% ggplot(aes(x=n)) +
  geom_histogram(binwidth=10) + xlab("Number of playlist additions per track") + 
  ylab("Number of tracks")

# Percentage of tracks added to more than 20 playlists
data %>% count(track_id) %>% filter(n>10) %>% nrow()/nrow(distinct_tracks)
```

Most listed tracks
```{r}
data %>% filter(!is.na(first_artistname)) %>% count(track_id, track_name, first_artistname, all_artistname) %>% 
  arrange(desc(n))
```

Major label tracks and indie label tracks
```{r}
# Number and percentage of tracks in data
distinct_tracks %>% count(major_label) %>% mutate(perc = n/nrow(distinct_tracks))
data %>% count(major_label)

# Number of playlist additions per track
data %>% group_by(major_label) %>% count(track_id) %>% summarise(mean(n), sd(n), min(n), max(n))
```

### 4.3.1 Playlists

Number of distinct playlists in data
```{r}
distinct_playlists <- data %>% distinct(pl_id, .keep_all=TRUE)
nrow(distinct_playlists)

# Additions per playlist
data %>% count(pl_id) %>% summarise(mean(n), sd(n), min(n), max(n))
```

Plot number of playlists per genre
```{r}
genre_counts <- distinct_playlists %>% summarise(pop=sum(pop), edm=sum(edm), tropicalhouse=sum(tropicalhouse), dancepop=sum(dancepop), rap=sum(rap), modernrock=sum(modernrock), electrohouse=sum(electrohouse), poprap=sum(poprap), 
    indierb=sum(indierb), indietronica=sum(indietronica), indiefolk=sum(indiefolk), postteenpop=sum(postteenpop), indierock=sum(indierock), trapmusic=sum(trapmusic), house=sum(house), latin=sum(latin), hiphop=sum(hiphop), indiepop=sum(indiepop), rock=sum(rock),  reggaeton=sum(reggaeton))

genre_counts_2 <- data.table(matrix(ncol=2, nrow=20))
colnames(genre_counts_2) <- c("genre", "frequency")
genre_counts_2$genre <- c("Pop", "EDM", "Tropical house", "Dance pop", 'Rap', 'Modern rock', 'Electro house', 'Pop rap', 'Indie R&B', 'Indietronica', 'Indie folk', 'Post-teen pop', 'Indie rock', 'Trap music', 'House', 'Latin', 'Hip hop', 'Indie pop', 'Rock', 'Reggaeton')
for (row in 1:nrow(genre_counts_2)) {
  genre_counts_2$frequency[row] <- genre_counts[row]
}
genre_counts_2$frequency <- as.numeric(genre_counts_2$frequency)

genre_counts_2 %>% ggplot(aes(x=reorder(genre, -frequency), y=frequency)) +
  geom_bar(stat="identity") + ylab("Frequency") + 
  geom_text(stat='identity', aes(label=frequency), vjust=-.5, size = 3.5) +
  theme(axis.text.x = element_text(angle = 90)) + xlab("Genre") + 
  coord_cartesian(ylim=c(0,37000))
```

Playlist owners
```{r}
distinct_playlists %>% count(owner) %>% mutate(perc=n/nrow(distinct_playlists))

data %>% count(owner)
data %>% group_by(owner) %>% count(pl_id) %>% summarise(mean(n), sd(n))
```

Playlist followers
```{r}
# mean and max n pl followers:
distinct_playlists %>% summarise(mean(pl_followers), sd(pl_followers), max(pl_followers))

distinct_playlists %>% group_by(owner) %>% summarise(mean(pl_followers), sd(pl_followers), median(pl_followers),
                                                     min(pl_followers), max(pl_followers))
distinct_playlists %>% group_by(owner) %>% summarise(foll=sum(pl_followers)) %>% mutate(perc=foll/sum(foll))

```


Plot frequency of major vs. indie additions per playlist owner
```{r}
# Major label
data %>% filter(owner=="Major label") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2)) %>% 
  ggplot(aes(x=major_label, y = perc)) +
  geom_bar(stat="identity") +
  xlab("Track label") + ylab("Percentage of playlist additions") +
  geom_text(stat="identity", aes(label=perc), vjust=-.5, size = 3.5) +
  ggtitle("Major label playlists") +
  coord_cartesian(ylim=c(0,100))

# Spotify
data %>% filter(owner=="Spotify") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2)) %>% 
  ggplot(aes(x=major_label, y = perc)) +
  geom_bar(stat="identity") +
  xlab("Track label") + ylab("Percentage of playlist additions") +
  geom_text(stat="identity", aes(label=perc), vjust=-.5, size = 3.5) +
  ggtitle("Official Spotify playlists") +
  coord_cartesian(ylim=c(0,100))

data %>% filter(owner=="Spotify", pop_above_median=="Unpopular") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2))
data %>% filter(owner=="Spotify", pop_above_median=="Popular") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2))

# Independent
data %>% filter(owner=="Other") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2)) %>% 
  ggplot(aes(x=major_label, y = perc)) +
  geom_bar(stat="identity") +
  xlab("Track label") + ylab("Percentage of playlist additions") +
  geom_text(stat="identity", aes(label=perc), vjust=-.5, size = 3.5) +
  ggtitle("Independent playlists") +
  coord_cartesian(ylim=c(0,100))

data %>% filter(owner=="Other", pop_above_median=="Unpopular") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2))
data %>% filter(owner=="Other", pop_above_median=="Popular") %>% 
  count(major_label) %>% mutate(perc=round(n/sum(n)*100, digits=2))
```
